---
title: "Tables"
output: bookdown::pdf_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(kableExtra)
library(tidyverse)
library(labelled)
library(ggplot2)
library(gridExtra)
```

# Visit event count

```{r}
events_data <- read.csv('./data/events_data_2020-07-27.csv', header = TRUE, na.strings = '', stringsAsFactors = FALSE, colClasses = c('univ_id' = 'character', 'univ_id_req' = 'character', 'school_id' = 'character', 'event_type' = 'character')) %>% as_tibble()
univ_info <- read.csv('./data/univ_data.csv', header = TRUE, na.strings = '', stringsAsFactors = FALSE, colClasses = c('univ_id' = 'character', 'zip_code' = 'character')) %>% as_tibble()

df <- events_data %>%
  select(univ_id, event_loc, event_type) %>% 
  left_join(univ_info %>% select(univ_id, univ_name), by = 'univ_id') %>% 
  mutate(event_loc = recode(event_loc, 'instate' = 'in', 'outofstate' = 'out'),
         event_type = recode(event_type, 'pub_hs' = 'pub', 'priv_hs' = 'priv', 'pub_4yr' = 'other', 'pse' = 'other'),
         event_cat = str_c(event_type, '_', event_loc),
         univ_name = recode(univ_name,
                            'The University of Alabama' = 'University of Alabama',
                            'Tulane University of Louisiana' = 'Tulane University',
                            'North Carolina State University at Raleigh' = 'North Carolina State University',
                            'University of Colorado Boulder' = 'University of Colorado-Boulder',
                            'University of Cincinnati-Main Campus' = 'University of Cincinnati',
                            'University of Pittsburgh-Pittsburgh Campus' = 'University of Pittsburgh',
                            'Sewanee-The University of the South' = 'Sewanee'
                            )) %>% 
  select(univ_id, univ_name, event_cat) %>% 
  group_by(univ_id, univ_name, event_cat) %>% 
  summarise(count = n()) %>% 
  pivot_wider(names_from = event_cat, values_from = count, values_fill = list(count = 0)) %>% 
  mutate(all_in = pub_in + priv_in + cc_in + other_in,
         all_out = pub_out + priv_out + cc_out + other_out,
         other_out = cc_out + other_out,
         all = all_in + all_out) %>% 
  arrange(univ_name)

pub_df <- df %>% filter(univ_id %in% (univ_info %>% filter(classification == 'public_research'))$univ_id)
priv_df <- df %>% filter(univ_id %in% (univ_info %>% filter(classification != 'public_research'))$univ_id, univ_id != '168218')

write.csv(pub_df, './data/event_count_pub.csv', row.names = FALSE)
write.csv(priv_df, './data/event_count_priv.csv', row.names = FALSE)
```

# Private HS universe

```{r}
privhs_universe <- readRDS('./data/privhs_universe.RDS')

# Filter for private HS that fits our criteria
# https://github.com/ksalazar3/recruiting-static-visuals/blob/62b08a202ca7c93433a91d7e72e6c0e93ac6eb5b/sql/update_imported_hs_data.sql#L52-L56
val_labels(privhs_universe$school_type)
privhs_universe %>% group_by(year) %>% count()

v_get_val <- Vectorize(function(x) str_to_lower(val_label(privhs_universe$region, x)))

df <- privhs_universe %>%
  filter(school_type %in% c(1, 2, 5), total_12 > 10) %>% 
  mutate(
    region = v_get_val(region),
    pct_blacklatinxnative = pct_black + pct_hispanic + pct_amerindian + pct_nativehawaii,
    pct_blacklatinxnative_cat = case_when(
      pct_blacklatinxnative < 10 ~ 'c1_lt10',
      pct_blacklatinxnative < 25 ~ 'c2_10to25',
      pct_blacklatinxnative < 50 ~ 'c3_25to50',
      pct_blacklatinxnative >= 50 ~ 'c4_50+'
    ),
    rank_cat2 = case_when(
      rank_within_category <= 200 & overall_niche_letter_grade == 'A+' ~ 'c1_top200',
      rank_within_category > 200 & overall_niche_letter_grade == 'A+' ~ 'c2_A+',
      overall_niche_letter_grade == 'A' ~ 'c3_A',
      !is.na(overall_niche_letter_grade) ~ 'c4_ltA'
    ) 
  ) %>% 
  select(total_12, region, religion, pct_blacklatinxnative_cat, rank_cat2)

plot_characteristic <- function(var_name, type, characteristic, label_text) {

  sub_df <- df %>% group_by(get(var_name)) %>% 
    summarize(unweighted = n(), weighted = sum(total_12)) %>% 
    pivot_longer(cols = c('unweighted', 'weighted'),
                 names_to = 'universe',
                 values_to = 'value')
  names(sub_df) <- c('key', 'universe', 'value')
  sub_df <- sub_df %>% filter(!is.na(key))
  
  sub_df$universe <- factor(sub_df$universe, levels = c('weighted', 'unweighted'))
  sub_df$key <- factor(sub_df$key, levels = characteristic)
  
  ggplot(sub_df, aes(fill = key, y = value, x = universe)) + 
      geom_bar(stat='identity', position = position_fill(reverse = TRUE), width = 0.7) +
      ggtitle('Title') +
      xlab('') + ylab('') +
      labs(fill = type) +
      scale_fill_manual(labels = label_text, values = c('#F8766D', '#7CAE00', '#00BFC4', '#C77CFF')) +
      scale_y_continuous(labels = scales::percent, expand = c(0, 0, 0.05, 0)) +
      theme(legend.position = 'top',
            legend.title = element_text(size = 6, face = 'bold'),
            legend.text = element_text(size = 6),
            legend.margin = margin(0, 0, 0, 0),
            legend.box.margin = margin(0, 0, -5, -30),
            legend.key.size = unit(0.3, 'cm'),
            panel.background = element_blank(),
            axis.ticks.y = element_blank(),
            text = element_text(size = 8),
            plot.title = element_text(color = 'white')) +
      coord_flip()
}

pdf(file.path('.', 'assets', 'figures', 'plot_pss_universe.pdf'), width = 8, height = 4)
par(mar=c(0, 0, 0, 0) + 0.1, mai=c(0, 0, 0, 0))
par(mfrow=c(2, 2))

a <- plot_characteristic('region', 'Region', c('northeast', 'midwest', 'south', 'west'), c('Northeast', 'Midwest', 'South', 'West'))
b <- plot_characteristic('religion', 'Religion', c('catholic', 'christian', 'other', 'nonsectarian'), c('Catholic', 'Christian', 'Other religion', 'Nonsectarian'))
c <- plot_characteristic('pct_blacklatinxnative_cat', '% Black/Latinx/Native', c('c1_lt10', 'c2_10to25', 'c3_25to50', 'c4_50+'), c('<10%', '10-25%', '25-50%', '50%+'))
d <- plot_characteristic('rank_cat2', 'Ranking', c('c1_top200', 'c2_A+', 'c3_A', 'c4_ltA'), c('Top 200 A+', 'Other A+', 'A', 'Less than A'))

grid.arrange(a, b, c, d, nrow=2, ncol=2)
dev.off()
```


# Top visited private HS (2-mode network)

```{r}
df <- readRDS('./assets/tables/table_2mode.RDS')

knitr::kable(head(df, n = 20)) %>% kable_styling(latex_options = 'scale_down')
```

Table \@ref(tab:hs-print-privu) shows characteristics of the 20 private high schools that were visited by the most private colleges and universities

```{r, hs-print-privu, echo = FALSE}
df <- readRDS('./assets/tables/table_2mode.RDS')

knitr::kable(head(df, n = 20, caption = 'ego table 3')) %>% kable_styling(latex_options = 'scale_down')
```

# Aggregated visited private HS (2-mode network)

```{r}
df <- readRDS('./assets/tables/table_2mode_agg.RDS')

knitr::kable(df) %>% kable_styling(latex_options = 'scale_down')
```

# Ego network table

```{r}
df <- readRDS('./assets/tables/table_ego.RDS')
knitr::kable(df %>% select(-ID, -Characteristics), row.names = F) %>% kable_styling(latex_options = 'scale_down')
```

# Ego network barplot

\newpage

```{r, fig.width=3.5, fig.height=3, message=FALSE}
# ggplot's default color palette
# library(scales)
# show_col(hue_pal()(4))

plot_characteristic <- function(df, type, characteristic, label_text, visit_count = FALSE) {
  if (visit_count) {
    df <- df %>% mutate(univ_name = str_c(univ_name, ' (N=', total, ')'))
  }
  
  ordered_df <- df %>% arrange(desc(cluster), univ_name)
  univ_order <- ordered_df$univ_name
  cluster_face <- map_chr(ordered_df$cluster, function(x) if_else(as.numeric(x) %% 2 == 0, 'bold', 'plain'))
  
  sub_df <- df %>% select(univ_name, all_of(characteristic)) %>%
    gather(key, value, characteristic) %>%
    mutate(value = parse_number(value))

  sub_df$univ_name <- factor(sub_df$univ_name, levels = rev(univ_order))
  sub_df$key <- factor(sub_df$key, levels = characteristic)

  ggplot(sub_df, aes(fill = key, y = value, x = univ_name)) + 
    geom_bar(stat='identity', position = position_fill(reverse = TRUE)) +
    ggtitle('Title') +
    xlab('') + ylab('') +
    labs(fill = type) +
    scale_fill_manual(labels = label_text, values = c('#F8766D', '#7CAE00', '#00BFC4', '#C77CFF')) +
    scale_y_continuous(labels = scales::percent, expand = c(0, 0, 0.05, 0)) +
    theme(legend.position = 'top',
          legend.title = element_text(size = 6, face = 'bold'),
          legend.text = element_text(size = 6),
          legend.margin = margin(0, 0, 0, 0),
          legend.box.margin = margin(0, 0, -5, -30),
          legend.key.size = unit(0.3, 'cm'),
          panel.background = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.y = element_text(face = rev(cluster_face)),
          text = element_text(size = 8),
          plot.title = element_text(color = 'white')) +
    coord_flip()
}


type_vec <- c('privu', 'pubu', 'pubu_outst', 'all', 'all_pubu_outst')

for (i in 1:length(type_vec)) {
  ego_df <-readRDS(file.path('.', 'assets', 'tables', str_c('table_ego_', type_vec[i], '.RDS')))

  pdf(file.path('.', 'assets', 'figures', str_c('ego_network_figure_', type_vec[i], '.pdf')))
  par(mar=c(0, 0, 0, 0) + 0.1, mai=c(0, 0, 0, 0))
  par(mfrow=c(2, 2))
  
  a <- plot_characteristic(ego_df, 'Region', c('northeast', 'midwest', 'south', 'west'), c('Northeast', 'Midwest', 'South', 'West'), visit_count = TRUE)
  b <- plot_characteristic(ego_df, 'Religion', c('catholic', 'christian', 'other', 'nonsectarian'), c('Catholic', 'Christian', 'Other religion', 'Nonsectarian'))
  c <- plot_characteristic(ego_df, '% Black/Latinx/Native', c('c1_lt10', 'c2_10to25', 'c3_25to50', 'c4_50+'), c('<10%', '10-25%', '25-50%', '50%+'), visit_count = TRUE)
  d <- plot_characteristic(ego_df, 'Ranking', c('c1_top200', 'c2_A+', 'c3_A', 'c4_ltA'), c('Top 200 A+', 'Other A+', 'A', 'Less than A'))
  
  grid.arrange(a, b, c, d, nrow=2, ncol=2, widths=c(1, 0.92))
  dev.off()
  
  # Need to use print() for ggplot inside loop in order for it to appear
  # https://stackoverflow.com/a/15678596/6373540
  # print(a)
  # print(b)
  # print(c)
  # print(d)
}
```

```{r ego-network-figure-privu, fig.align = 'center', echo = FALSE, fig.cap = "Characteristics of private high schools visited by private colleges and universities"}
knitr::include_graphics('assets/figures/ego_network_figure_privu.pdf')
```

\begin{center}
* \textit{Universities are grouped by clusters, as indicated by alternating bolded and normal font}
\end{center}

